---
title: "Sustainable food systems\nAnalysis"
author: "Harold Achicanoy & Patricia Alvarez"
date: "3/15/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data we use

For test purposes we define some theoretical dimensions could influence Sustainable Food Systems. These dimensions were:

- **Nutritional aspects (NUTR)**
- **Human interventions (HINT)**
- **Food security (FSCY)**

On the other hand, **Sustainable Food Systems (SUFS)** is defined as the objective dimension.

## Nutritional aspects (NUTR)

Uses indicators of:

- Prevalence of Child Malnutrition (1990 - 2002)
- Global hunger index (2016)

## Human interventions (HINT)

Uses indicators of:

- Accessibility travel time to major cities
- Human Influence Index (HII) normalized by biome and realm

## Food security (FSCY)

Uses indicators of:

- Access to improved sanitation facilities (%)
- Access to improved water sources (%)
- Gross domestic product per capita, PPP (constant 2011 international $)
- Political stability and absence of violence/terrorism (index)

## PLS-PM in action

Here it's very important to have or develop a theoretical model that link dimensions and take under consideration the relations between them.

```{r, include = TRUE}
suppressMessages(library(plspm))
NUTR <- c(0, 0, 0, 0)
HINT <- c(0, 0, 0, 0)
FSCY <- c(0, 0, 0, 0)
SUFS <- c(1, 1, 1, 0)
sfs_path <- rbind(NUTR, HINT, FSCY, SUFS); rm(NUTR, HINT, FSCY, SUFS)
colnames(sfs_path) <- rownames(sfs_path)
plot(sfs_path)
```

## Results

```{r cars, echo = TRUE}
summary(cars)
```

## Following steps

```{r pressure}
plot(pressure)
```

