---
title: "Sustainable food systems\nPreliminary analysis"
author: "Harold Achicanoy, Patricia Alvarez and Lea Lamotte"
date: "3/10/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data we use

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Missing values and what we did with them ...

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## PLS-PM in action

- Bullet 1
- Bullet 2
- Bullet 3

## Results

```{r cars, echo = TRUE}
summary(cars)
```

## Following steps

```{r pressure}
plot(pressure)
```

